\textbf{Dennis Muiruri} is a member of the empirical software engineering research group at the University of Helsinki. His \DIFdelbegin \DIFdel{current }\DIFdelend research interests include deployment and operations of \DIFdelbegin \DIFdel{machine learning }\DIFdelend \DIFaddbegin \DIFadd{ML }\DIFaddend systems.

\textbf{Lucy Ellen Lwakatare} is a postdoc at the University of Helsinki. She received her PhD from University of Oulu\DIFdelbegin \DIFdel{in 2013. }\DIFdelend \DIFaddbegin \DIFadd{.
%DIF > in 2013. 
}\DIFaddend Her research interests include \DIFdelbegin \DIFdel{: Agile}\DIFdelend \DIFaddbegin \DIFadd{agile}\DIFaddend , DevOps and ML engineering.

\textbf{Jukka K. Nurminen} is a professor at the University of Helsinki. 
\DIFdelbegin \DIFdel{He has worked extensively on software research in industry at Nokia Research Center, in academia at Aalto University, and in applied research at VTT. He }\DIFdelend %DIF > He has worked extensively on software research in industry at Nokia Research Center, in academia at Aalto University, and in applied research at VTT. 
\DIFaddbegin \DIFadd{He }\DIFaddend received PhD from Helsinki University of Technology\DIFdelbegin \DIFdel{(now Aalto University)}\DIFdelend . His main interests are on tools and techniques for \DIFdelbegin \DIFdel{developing }\DIFdelend data-intensive software systems\DIFdelbegin \DIFdel{including }\DIFdelend \DIFaddbegin \DIFadd{, }\DIFaddend testing of AI solutions, \DIFdelbegin \DIFdel{computational moral }\DIFdelend and software development for quantum computers.

\textbf{Tommi Mikkonen} \DIFaddbegin \DIFadd{is a professor of software engineering at University of Helsinki and University of Jyv\"{a}skyl\"{a}. He }\DIFaddend received his Ph.D. from Tampere University of Technology\DIFdelbegin \DIFdel{year 1999. He is a full professor in University of Helsinki of the empirical software engineering research group. His research }\DIFdelend \DIFaddbegin \DIFadd{. His }\DIFaddend interests include IoT, software architectures, and software engineering for AI.
